### 2019-04-23 2-Chapters/1-Chapter/chapter1.tex
Talk to Emilie

\begin{itemize}
\item
    \emph{GNU Radio Implementation of MALIN: ``Multi-Armed bandits Learning for Internet-of-things Networks''},
    by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
    \texttt{HAL.Inria.fr/hal-02006825}.
    \cite{Besson2019WCNC}

\item
    \emph{Multi-Player Bandits Revisited},
    by \textbf{Lilian Besson} \& Emilie Kaufmann.
    \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01629733}.
    \cite{Besson2018ALT}

\item
    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01705292}.
    \cite{Besson2018WCNC}

\item
    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
    \texttt{HAL.Inria.fr/hal-01575419},
    \textbf{Best Paper Award}.
    \cite{Bonnefoi17}

@@

\item
    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
    1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019.
    \cite{Bonnefoi2019WCNC}

\item
    \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
    by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
    Demonstration presented at \emph{International Conference on Communication} (ICT, \texttt{ict-2018.org/demos}), in Saint-Malo, France in June $2018$.
    See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.
    \cite{Besson2018ICT}

@@

\begin{itemize}
\item
    \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
    submitted at GRETSI 2019, \texttt{GRETSI.fr/colloque2019},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.
    \cite{Besson2019Gretsi}

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    submitted at COLT 2019, \texttt{LearningTheory.org/colt2019/},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.
    \cite{Besson2019GLRT}

@@

\begin{itemize}
\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson}, active development since October $2016$,
    \texttt{HAL.Inria.fr/hal-01840022}.
    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
    \cite{SMPyBandits,SMPyBanditsJMLR}

\item
    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
    \texttt{HAL.Inria.fr/hal-01736357}.
    \cite{Besson2018DoublingTricks}

@@

\begin{itemize}
\item
    \emph{A Note on the Ei Function and a Useful Sum-Inequality},
    by \textbf{Lilian Besson},
    February $2018$,
    \texttt{HAL.Inria.fr/hal-01847480}.

@@

\paragraph{Responsabilities}
    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf. \texttt{HAL.Inria.fr/hal-02013839/document}}.
    I also helped for the organization and presented another poster\footnote{Cf. \texttt{HAL.Inria.fr/hal-02013847/document}} at IETR seminar (\texttt{seminar-ietr-18.sciencesconf.org}), held in Vannes in June $2018$.

\paragraph{Reviews}
	for \emph{European Workshop on Reinforcement Learning} (Lille) \texttt{ewrl.wordpress.com/ewrl14-2018}, in October $2018$, I reviewed $5$ papers.
	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

>>>>

\begin{itemize}
\item
    \emph{GNU Radio Implementation of MALIN: ``Multi-Armed bandits Learning for Internet-of-things Networks''},
    by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019,
    \href{https://HAL.Inria.fr/hal-02006825}{\texttt{HAL.Inria.fr/hal-02006825}}.
    \cite{Besson2019WCNC}

\item
    \emph{Multi-Player Bandits Revisited},
    by \textbf{Lilian Besson} \& Emilie Kaufmann.
    \emph{Algorithmic Learning Theory} (ALT, \href{http://www.cs.cornell.edu/conferences/alt2018}{\texttt{www.cs.cornell.edu/conferences/alt2018}}), Lanzarote, Spain, April 2018,
    \href{https://HAL.Inria.fr/hal-01629733}{\texttt{HAL.Inria.fr/hal-01629733}}.
    \cite{Besson2018ALT}

\item
    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2018.ieee-wcnc.org}{\texttt{wcnc2018.ieee-wcnc.org}}), Barcelone, Spain, April 2018,
    \href{https://HAL.Inria.fr/hal-01705292}{\texttt{HAL.Inria.fr/hal-01705292}}.
    \cite{Besson2018WCNC}

\item
    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \href{http://crowncom.org/2017}{\texttt{crowncom.org/2017}}), Lisboa, Portugal, Septembre $2017$,
    \href{https://HAL.Inria.fr/hal-01575419}{\texttt{HAL.Inria.fr/hal-01575419}},
    \textbf{Best Paper Award}.
    \cite{Bonnefoi17}

@@

\item
    \emph{Decentralized Spectrum Learning for IoT Wireless Networks Collision Mitigation},\\
    by Christophe Moy \& \textbf{Lilian Besson}.
    1st International ISIoT workshop (\href{https://sites.google.com/view/ISIoT2019}{\texttt{sites.google.com/view/ISIoT2019}}),
    at \emph{Conference on Distributed Computing in Sensor Systems} (IEEE DCOSS, \href{http://2019.dcoss.org}{\texttt{2019.dcoss.org}}), Santorini, Greece, May 2019.
    \cite{MoyBesson2019}

\item
    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
    1st International MOTIoN workshop (\href{https://sites.google.com/view/wcncworkshop-motion2019}{\texttt{sites.google.com/view/wcncworkshop-motion2019}}),
    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \href{http://wcnc2019.ieee-wcnc.org}{\texttt{wcnc2019.ieee-wcnc.org}}), Marrakech, Morocco, April 2019.
    \cite{Bonnefoi2019WCNC}

\item
    \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
    by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
    Demonstration presented at \emph{International Conference on Communication} (ICT, \href{http://ict-2018.org/demos}{\texttt{ict-2018.org/demos}}), in Saint-Malo, France in June $2018$.
    See \href{https://YouTu.be/HospLNQhcMk}{\texttt{YouTu.be/HospLNQhcMk}} for a $6$-minutes presentation video.
    \cite{Besson2018ICT}

@@

\begin{itemize}
\item
    \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
    submitted at GRETSI 2019, \href{http://GRETSI.fr/colloque2019}{\texttt{GRETSI.fr/colloque2019}},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
    \cite{Besson2019Gretsi}

@@

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
    \href{https://HAL.Inria.fr/hal-02006471}{\texttt{HAL.Inria.fr/hal-02006471}}.
    \cite{Besson2019GLRT}

\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson}, active development since October $2016$,
    \href{https://HAL.Inria.fr/hal-01840022}{\texttt{HAL.Inria.fr/hal-01840022}}.
    About $40000$ lines of code, hosted on \href{https://GitHub.com/SMPyBandits}{\texttt{GitHub.com/SMPyBandits}},
    with a complete documentation on \href{https://SMPyBandits.rtfd.io}{\texttt{SMPyBandits.rtfd.io}}.
    \cite{SMPyBandits,SMPyBanditsJMLR}

\item
    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
    \href{https://HAL.Inria.fr/hal-01736357}{\texttt{HAL.Inria.fr/hal-01736357}}.
    \cite{Besson2018DoublingTricks}

@@

\begin{itemize}
\item
    \emph{A Note on the Ei Function and a Useful Sum-Inequality},
    by \textbf{Lilian Besson},
    February $2018$,
    \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.

@@

\paragraph{Responsabilities}
    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{Cf. \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
    I also helped for the organization and presented another poster\footnote{Cf. \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Evaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.

\paragraph{Reviews}
	for \emph{European Workshop on Reinforcement Learning} (Lille) \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}, in October $2018$, I reviewed $5$ papers.
	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

### 2019-04-23 2-Chapters/3-Chapter/chapter3.tex
Talk to Emilie

In this Chapter~\ref{chapter:3}, we start by presenting in Section~\ref{sec:3:presentationLibrary} the organization of the library.
We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms. We focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling.
% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
% but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
%
Finally, in Appendix~\ref{sec:3:appendix} we include some minimalist files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.

@@

@@

\subsubsection{Features}

@@

\subsubsection{How to run experiments?}

@@

\subsection{Continuous integration using Read the Docs and Travis CI}

% FIXME
\TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to:

1. automatically build the documentation on Read the Docs after every commit. This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.

2. automatically run short numerical simulations on Travis CI after every commit. This allows to check that no modification on any part of the codebase breaks anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proved very useful to quickly find and fix new bugs.
}

@@

$(i)$
Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic (see Chapter~\ref{chapter:6}) problems of single- or multi-player multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also compute, store and visualize other measures such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

@@

% WARNING
% \paragraph{Remark.}
We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered one of the main contribution of this thesis.
%
Our library is used for the numerical simulations in all this thesis, except Chapter~\ref{chapter:4}. We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in Appendix~\ref{app:2:DoublingTricks}.

\paragraph{Future works.}
%
An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-player bandit model,
and even if some have already been implemented, a major future work is to implement the most interesting ones
(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).

Since $2017$, I also received by email or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
Its documentation has seen already about $20000$ visits, the project on GitHub had $74$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
%
We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.

As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
I would also love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.

@@

We present below in Listing~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
which is used to define the experiment.
It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
It uses environment variables, so a user can configure some parameters of the simulation when launching them:

>>>>

In this Chapter~\ref{chapter:3}, we start by presenting in Section~\ref{sec:3:presentationLibrary} the organization of the library.
We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
then we present in Section~\ref{sec:3:timeAndMemoryCosts} additional numerical simulations to discuss about the time efficient and memory costs of the some MAB algorithms. We focus on the ones being used in the rest of this thesis, \UCB, \klUCB, and Thompson Sampling.
% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
% but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
%
Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, as well as details regarding the use of parallel computing to speed-up simulations.

@@

We have not yet tried to add support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
%
However, note that our library does support finite-state real-valued Markov MAB models, where arms do not correspond to distributions but to Markov chains, in the two variants of \emph{rested} or \emph{restless} Markov chains, as introduced by \cite{Anantharam87b}.

@@

\subsection{Features}

@@

\subsection{How to run experiments?}

@@

\subsection{Continuous integration with Read the Docs and Travis CI}

% \TODOL{I need to quickly explain that SMPyBandits is using Read the Docs since $2017$ and Travis Ci since $2018$ to: }

Since $2017$ and $2018$, SMPyBandits is using the free version of two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
%
These CI services are free for open-source work, and are both triggered whenever a modification of the codebase is sent to the hosting platform (\ie, GitHub).

\paragraph{Read the Docs.}
Since $2017$, we have been using the free web service provided by Read the Docs (\href{https://readthedocs.org/}{\texttt{ReadTheDocs.org}}).
Read the Docs allows to automatically build the documentation after every commit. This allows to regularly check that the library is well formatted and can be imported correctly, as well as keeping the online documentation up-to-date.
Moreover, they offer to host the documentation online, at \href{https://smpybandits.rtfd.io/}{\texttt{SMPyBandits.ReadTheDocs.io}}.
For more details, see \href{https://readthedocs.org/projects/smpybandits/}{\texttt{https://ReadTheDocs.org/projects/SMPyBandits}}.
%
Since its first use, the service did about 180 builds, and about $10\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.readthedocs.yml} file in SMPyBandits main folder, and it builds the documentation in about two minutes.
I have also been building the documentation manually on a weekly basis, to also host it on \href{https://smpybandits.github.io/}{\texttt{SMPyBandits.GitHub.io}} thanks to GitHub pages.
In the last two years, the documentation saw about 8500 unique visits, while the GitHub project have been visited about 3500 unique visits.

\paragraph{Travis CI.}
Since $2018$, we have been using the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
Travis CI allows to automatically run short numerical simulations after every commit.
This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proved very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-player model).
Builds typically run for $15$ minutes, and Travis CI have been proved to be very useful in our development process.

@@

$(i)$
Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
One of the main quality of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

@@

% WARNING
% \paragraph{Remark.}
We want to conclude by highlighting that a significant amount of time during my PhD was devoted to the development of SMPyBandits, and as such the library, its documentation and this chapter  are considered an important contribution of this thesis.
%
Our library is used for the numerical simulations in all this thesis, except Chapter~\ref{chapter:4}.
% We use it to compare state-of-the-art single-player algorithms designed for stationary MAB in Chapter~\ref{chapter:2}, for multi-armed problems in Chapter~\ref{chapter:5} and for piece-wise stationary problems in Chapter~\ref{chapter:6}.
We also used it in other publications not included in this thesis, like our work on doubling tricks \cite{Besson2018DoublingTricks} that we shortly present in Appendix~\ref{app:2:DoublingTricks}.

\paragraph{Future works.}
%
An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-player bandit model,
and even if some have already been implemented, a major future work is to implement the most interesting ones
(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).\\
%
\indent
Since $2017$, I also received by email or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
Its documentation has seen already about $20000$ visits, the project on GitHub had $76$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $15$ to $25$ researchers in other labs we use or used SMPyBandits.
While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
%
We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope to be accepted, as it would give more visibility to our work.\\
%
\indent
As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
I would also love to be able in the next to teach a graduate or under-graduate course on multi-armed bandits, and use my library as a support to illustrate the course, as well as practical sessions.

@@

We present below in Listing~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
which is used to define the experiment.
It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:

### 2019-04-23 2-Chapters/4-Chapter/chapter4.tex
Go
Talk to Emilie

\paragraph{Abstract}

@@

Possible future extensions of our demonstration include the following points.
We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card is enough to simulate another gateway).
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

>>>>

\newpage
\paragraph{Abstract}

@@

Possible future extensions of our demonstration include the following points.
We could consider more dynamic devices (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more devices with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card to simulate another gateway).
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.
